---
title: "Processing Google Takeout Fitbit Data"
author: Genevieve Roberts
output:
  pdf_document:
    latex_engine: pdflatex  # Specify the LaTeX engine
date: "2025-08-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(warn = -1)
library(knitr)
library(kableExtra)
library(pander)
```

## Load Packages and Setup Functions and Constants

```{r packages}
library(here)
library(tidyverse)
```

```{r functions}
source(here::here("fitbit/takeout_fitbit_processing_functions.R"))
source(here::here("fitbit/fitbit_plotting_functions.R"))
```

```{r data path}
#current path
data_path <- here::here("fitbit/sample_fitbit_takeout_data/9Aug25_groberts_fitbit_takeout/Fitbit")
print(data_path)
```
\newpage
# Explore some FitBit data
```{r constants}
#define some constants for the nb
start_date="2021-01-07"
end_date="2025-08-09"

#add some dates of interest to highlight
dates_of_interest_start = "2025-07-29"
dates_of_interest_end = "2025-08-01"
```

## Heart Rate Variablility
```{r load hrv data}
# Combined detailed + summary
hrv_data <- load_fitbit_hrv(start_date = start_date,
                            end_date = end_date, 
                            root_dir = data_path,
                            summary_only = FALSE)

# Only summary data
hrv_data_only <- load_fitbit_hrv(start_date = start_date,
                                end_date = end_date,
                               root_dir = data_path,
                               summary_only = TRUE)

pander(sample_n(hrv_data, 5))
```

\newpage
Here, I want to know if the `rmssd_summary` from the summary HRV files is simply
the mean across all of the "detail" datapoints for that day. The plot below
suggests they are not the same, but they are closely related.

```{r}
check_if_mean_equals_summary <- hrv_data %>%
  group_by(file_date) %>%
  summarize(
    mean_rmssd_detail = mean(rmssd_detail, na.rm = TRUE),
    rmssd_summary = first(rmssd_summary)  # summary has one value per date
  ) %>%
  ungroup()

# Prepare data in long format for plotting
mean_detail_compare_plot_df <- check_if_mean_equals_summary %>%
  pivot_longer(cols = c(mean_rmssd_detail, rmssd_summary),
               names_to = "Type",
               values_to = "RMSSD")

ggplot(mean_detail_compare_plot_df, aes(x = file_date, y = RMSSD, color = Type)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(
    title = "Comparison of RMSSD: Detail Mean vs. Summary",
    x = "Date",
    y = "RMSSD",
    color = "Measure"
  ) +
  theme_bw()
```
\newpage
The plot below shows how we can highlight a period of interest
```{r plot HRV with box around dates of interest}
# Example dates to highlight
highlight_start <- as.Date("2025-07-29")
highlight_end <- as.Date("2025-08-02")

#add the mean rmssd from the detailed HRV information
hrv_data <- hrv_data %>%
  group_by(file_date) %>%
  mutate(mean_rmssd_detail = mean(rmssd_detail)) %>%
  ungroup()

ggplot(hrv_data, aes(x = file_date, y = mean_rmssd_detail)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "blue", size = 2) +
  
  # Highlight date range with a transparent rectangle
  annotate(
    "rect",
    xmin = highlight_start, xmax = highlight_end,
    ymin = -Inf, ymax = Inf,
    alpha = 0.2, fill = "orange"
  ) +
  
  labs(
    title = "HRV Over Time with Highlighted Date Range",
    x = "Date",
    y = "HRV (Mean RMSSD Detail)"
  ) +
  
  theme_bw()
```
\newpage

## Add Resting Heart Rate Data

```{r combine hrv and rhr}
rhr_data <- load_fitbit_resting_hr(start_date = start_date,
                            end_date = end_date, 
                            root_dir = data_path)

# Combine by day (default)
combined <- combine_fitbit_data(hrv_data, rhr_data)
pander(sample_n(combined, 5))
```
\newpage

The plot below compares heart rate variability to resting heart rate.
You can see there is a rough inverse correlation.

```{r}
plot_dual_axis(
  data = combined,
  col1 = rmssd_detail,
  col2 = resting_hr,
  label1 = "HRV (RMSSD)",
  label2 = "Resting HR (bpm)",
  title = "Comparison of HRV and RHR",
  highlight_start = dates_of_interest_start,
  highlight_end = dates_of_interest_end
)
```
\newpage

### Daily Readiness Score
```{r}
# Load the daily readiness score and combine it with the other data
daily_ready <- load_fitbit_daily_readiness(start_date = start_date,
                            end_date = end_date, 
                            root_dir = data_path)

# Combine by day
combined <- combine_fitbit_data(combined, daily_ready)
```

The plot below shows a roughly inverse correlation between Daily Readiness Score
and Resting Heartrate:
```{r}
plot_dual_axis(
  data = combined,
  col1 = daily_readiness_score,
  col2 = resting_hr,
  label1 = "Daily Readiness Score \n(Higher is Better)",
  label2 = "Resting HR \n(bpm)",
  title = "Comparison of Daily Readiness Score\nand Resting Heartrate",
  highlight_start = dates_of_interest_start,
  highlight_end = dates_of_interest_end
)
```

